Flooding of the Okavango Delta influences Connectivity for Dispersing African Wild Dogs

### 2022-08-08 Chapter_8/04_Manuscript/Manuscript.tex
Various updates

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
Many ecosystems experience drastic changes in environmental conditions due to
seasonality. While such seasonal changes may drastically alter connectivity for
endangered species, most studies represent the environment by a static set of
spatial layers, thus ignoring seasonal changes. Here, we address this
shortcoming and employ individual-based simulations to investigate how seasonal
flooding of the Okavango delta influences connectivity for dispersing African
wild dogs (\textit{Lycaon pictus}). Our results show that the Okavango delta
poses a substantial dispersal barrier when the flood is at maximum extent, yet
that viable dispersal corridors exist when the flood is at a minimum level.
Despite a better understanding of the conservation needs for African wild dogs,
our study also provides evidence that incorporating seasonality in studies of
connectivity is imperative to more accurately predict dispersal ability of
endangered species.
\end{abstract}

@@

@@

\subsection{Flood}
In order to generate spatial layers that depict the flood extent at different
points in time, we ... \Cref{FloodExtent}.

@@

>>>>

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
Many ecosystems experience substantial changes in environmental conditions due
to seasonality. While such seasonal changes may drastically alter connectivity
for endangered species, most studies represent the environment by a static set
of spatial layers, thus ignoring seasonal variation. Here, we address this
shortcoming and employ individual-based simulations to investigate how seasonal
flooding of the Okavango delta influences connectivity for dispersing African
wild dogs (\textit{Lycaon pictus}). Our results show that the Okavango delta
poses a substantial dispersal barrier when the flood is at maximum extent, yet
that viable dispersal corridors exist when the flood is at a minimum level.
Despite a better understanding of the conservation needs for African wild dogs,
our study also provides evidence that incorporating seasonality in studies of
connectivity is imperative to more accurately predict the dispersal ability of
endangered species.
\end{abstract}

@@

However, \cite{Hofmann.2021} and \cite{Hofmann.2022} used a single floodmap...

@@

\subsection{Spatial Habitat Layers}
In order to generate spatial layers that depict the flood extent at different
points in time, we ... \Cref{FloodExtent}.

@@

\subsection{Dispersal Simulation}
We used a previously parametrized movement model to simulate dispersal of
African wild dogs. To simulate dispersing wild dogs, we employed the simulation
framework presented in \cite{Hofmann.2022}. In this framework, a movement model
that renders habitat and movement preferences of dispersing individuals is
parametrized using step-selection functions. Once parametrized, the model can be
employed to simulate virtual dispersers moving across the landscape.

We released virtual dispersers at random locations within the four distinct
source areas depicted in \Cref{StudyArea}. From each source area we simulated xx
individuals moving across the landscape at minimum flood and another xx
individuals at maximum flood. For comparison, we also ran the simulations for a
medium flood extent, yet the results for this will be presented in the appendix.

\subsection{Connectivity}
To gain insights into landscape connectivity from the simulated dispersal
events, we prepared three complementary connectivity maps (see again
\citep{Hofmann.2022}); a heatmap, indicating the absolute frequency at which
each pixel in the study area was traversed by a virtual disperser, a betweenness
map, highlighting pixels of particular importance to facilitate movement from
one area to another, and a map of interpatch connectivity, depicting the
presence and intensity of functional links between source areas. To compute the
heatmap, we counted how many times each pixel in the study area was traversed by
a virtual disperser. To calculate betweenness, we overlayed the study area with
a regular grid and used the simulated trajectories to determine movements from
one grid-cell to another. Specifically, we counted how often transitions from
one grid-cell to any other occured. The centerpoint of each grid-cell then
served as network node, and the number of transitions from one cell to another
as weights between the respective nodes. Based on the so generated network, we
then computed betweenness scores for all network nodes (i.e. for all grid-cells
across the study area). Lastly, to compute interpatch connectivity, we
determined the frequency at which virtual dispersers originating from one source
area reached any of the other source areas. We also calculated the average
dispersal duration needed until individuals reached the respective area.

\section{Results}

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Heatmaps2000.png}
  \caption{}
  \label{Heatmaps}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_DifferenceHeatmaps.png}
  \caption{}
  \label{DifferenceHeatmaps}
  \end{center}
\end{figure}

\section{Discussion}
